---
title: "CA Gas Forecase"
author: "ClareCallahan"
date: "4/11/2021"
output: html_document
---

Data was pulled on 4/11/2021 from the EIA https://www.eia.gov/dnav/pet/PET_PRI_GND_DCUS_SCA_M.htm
Values are in $/Gallon of averaged Regular Gas formulations in CA from Jan 1995-Mar 2021 

```{r}

#install.packages('forecast', dependencies = TRUE)

```



```{r}
library(ggplot2)
library(readr)
library(tidyverse)
library(lubridate)
library(forecast)
library(tseries)
library(dplyr)
library(mgcv)
library(astsa)

```

```{r}
temp <- tempfile()
download.file("http://download.geonames.org/export/zip/US.zip",temp)
con <- unz(temp, "US.txt")
US <- read.delim(con, header=FALSE)
unlink(temp)
colnames(US)[c(3,5,6)] <- c("city","state","county")
CA<-filter(US, state== "CA")
US$city <- tolower(US$city)

County<-(US)[c(3,6)]


#join on city from charging station 

station <- read_csv("alt_fuel_stations.csv")
station$city <- tolower(station$city)
data<-left_join(station, County, by = "city")



```






```{r}

data$Open_Date <-mdy(data$Open_Date)

#https://blogs.oracle.com/datascience/introduction-to-forecasting-with-arima-in-r
summary(data)
class(data$Open_Date)
data$Year <- year(data$Open_Date)

```



```{r}


data%>%
ggplot(aes(x=Open_Date, y=data$`EV Level2 EVSE Num`))+
  geom_line() + scale_x_date('month')+ 
    ylab("Count of Charging Stations Opened in Year X") +
    xlab("Date")


```


##creating counts for level 2 and DC by county by open data, cummulativly add prior years stations 

```{r}
data<-as.data.frame(data)

stationCounty <- data%>%
  group_by(county)%>%
  aggregate(data["Year"], by= list(year = Year), FUN= sum)



```












##Unnecessary step
```{r}
#clean and smooth 


gas_ts = ts(gas_price[, c('GasolinePrices')])

gas_price$clean_Gas = tsclean(gas_ts)


gas_price%>%
ggplot(aes(x=Date, y=clean_Gas))+
  geom_line() + scale_x_date('month')+ 
    ylab("GasPrice_smoothed") +
    xlab("Date")

  

```



#moving average 


gas_price$gas_ma = ma(gas_price$clean_Gas, order=7) # using the clean count with no outliers

gas_price$gas_ma30 = ma(gas_price$clean_Gas, order=30)



ggplot() + 
  geom_line(data = gas_price, aes(x = Date, y = clean_Gas, colour = " Price")) +
  geom_line(data = gas_price, aes(x = Date, y = gas_ma,   colour = "Weekly Moving Average"))  +
  geom_line(data = gas_price, aes(x = Date, y = gas_ma30, colour = "Monthly Moving Average"))  +
  ylab('Spot Price') 




```{r}

#decomposition 



count_ma = ts(na.omit(gas_price$GasolinePrices), frequency=12) 
decomp = stl(count_ma, s.window="periodic")
deseasonal_sp <- seasadj(decomp, allow.multiplicative.trend=TRUE)
plot(decomp)

plot(deseasonal_sp)
```

```{r}

count_ma %>%
  stl(s.window='periodic')%>%
  seasadj() -> eeadj
autoplot(eeadj)


```

##P-value >.05 Data is non-stationary
```{r}

adf.test(count_ma, alternative = "stationary")


#p-value >.05 indicates weak alternative hypothesis, i.e. do not reject the null that the data is non-stationary

```
```{r}
Acf(count_ma, main='')

Pacf(count_ma, main='')

#https://towardsdatascience.com/time-series-forecasting-arima-models-7f221e9eee06

```
##Data is now stationary 
```{r}


count_d1 = diff(deseasonal_sp, differences = 1)
plot(count_d1)
adf.test(count_d1, alternative = "stationary")


```



```{r}

Acf(count_d1, main='ACF for Differenced Series')
Pacf(count_d1, main='PACF for Differenced Series')

```


```{r}

fit<- auto.arima(deseasonal_sp, seasonal = FALSE)
tsdisplay(residuals(fit), lag.max = 2, main = 'Model Residuals')

checkresiduals(fit)
```


```{r}

fcast <- forecast(fit)
plot(fcast)


```







```{r}

plot(predict(deseasonal_sp))


prediction2 <- predict(fit[["x"]])

plot(prediction2)

```

```{r}

#132 months. Note predictions start April 2021 and continue to March 2035 (adding dates via excel)... after a certain point the values flatline 
futureVal <-predict(fit, n.ahead = 168)

#futurVal <- forecast(fit, level=c(95))

plot(futurVal)

gas_df<- as.data.frame(futureVal$pred)

#add in data back to 1995
write.csv(gas_df, "CA_GasPreds_2035.csv")
```