---
title: "Preliminary EDA for the BSF data"
author: "Arun/Daniel"
date: "2/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("ggplot2")
library("ggmap")
library("DescTools")
```

#Preliminary EDA

The notebook focusses on two aspects:

<b> 1. Typical purchaser profile
    2. Vehicle profile </b>

The notebook starts with reading in data. To carry out further analysis, the dates, amount etc were converted to either date type or to float type

Data shared by BSF

```{r}
cvap <- read.csv("CVAP.csv")
#cvap

## caution 12 geo coordinates missing
# cvap_lat <- read.csv("cvap longlat.csv")
# cvap_lat <- cvap_lat[,c("Name.Customer.ID", "Latitude", "Longitude")]

# cvap <- merge(cvap1, cvap_lat, by = "Name.Customer.ID")
```


```{r}
# paste("Total number of unique customers", nrow(cvap1 %>% group_by(Name.Customer.ID) %>% summarise(n =n())), sep = "  ")

paste("Total number of unique customers", nrow(cvap %>% group_by(Name.Customer.ID) %>% summarise(n =n())), sep = "  ")
```


## Data manipulation
```{r}
#clean the annual income
cvap$ANNUAL.INCOME <- sub(",","", cvap$ANNUAL.INCOME)
cvap$ANNUAL.INCOME <- sub("\\$","", cvap$ANNUAL.INCOME)
cvap$ANNUAL.INCOME <- as.numeric(cvap$ANNUAL.INCOME)
```

```{r}
#clean the buydown amount
cvap$BUYDOWN..AMOUNT <- sub(",","", cvap$BUYDOWN..AMOUNT)
cvap$BUYDOWN..AMOUNT <- sub("\\$","", cvap$BUYDOWN..AMOUNT)
cvap$BUYDOWN..AMOUNT <- as.numeric(cvap$BUYDOWN..AMOUNT)

# loan downpayment
cvap$Loan.down.payment <- sub(",","", cvap$Loan.down.payment)
cvap$Loan.down.payment <- sub("\\$","", cvap$Loan.down.payment)
cvap$Loan.down.payment <- as.numeric(cvap$Loan.down.payment)

#loan amount
cvap$LOAN.AMOUNT <- sub(",","", cvap$LOAN.AMOUNT)
cvap$LOAN.AMOUNT <- sub("\\$","", cvap$LOAN.AMOUNT)
cvap$LOAN.AMOUNT <- as.numeric(cvap$LOAN.AMOUNT)

#total vehical cost
cvap$TOTAL.VEHICLE.COST <- sub(",","", cvap$TOTAL.VEHICLE.COST)
cvap$TOTAL.VEHICLE.COST <- sub("\\$","", cvap$TOTAL.VEHICLE.COST)
cvap$TOTAL.VEHICLE.COST <- as.numeric(cvap$TOTAL.VEHICLE.COST)
```

## Purchaser's profile



### Buyer's economic status
```{r}

#economic status
cvap %>%
  ggplot(aes(x = DAC..Y.N., fill =DAC..Y.N.))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("DAC (Yes/No)")
```

### Buyer's Age profile 

```{r}
#age profile
cvap %>% 
  ggplot(aes(x = Age))+
  geom_density()+ ggtitle("Age Distribution")

ggplot(data=cvap, mapping=aes(sample=Age)) +
  stat_qq(geom = 'point', distribution = 'qunif') +
  scale_x_continuous(n.breaks = 10) +
  labs(x='Quantile', y='Age', title='Quantile Plot, Age')

```

### Buyer's tax status, loan source tapped, loan term sought

```{r}
#if filled taxes
cvap %>%
  ggplot(aes(x = Filed.Taxes.Last.Year, fill =Filed.Taxes.Last.Year))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("Tax filed last year")

#owns a home
cvap %>%
  ggplot(aes(x = Owns.Home, fill =Owns.Home))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("Owns Home?")

# loan source tapped into
cvap %>%
  ggplot(aes(x = Loan.Source, fill =Loan.Source))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("Loan Source")

# loan term
cvap %>%
  ggplot(aes(x = LOAN.TERM, fill =LOAN.TERM))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("Loan Source")


```


### Annual Income distribution

```{r}
# annual income distribution

cvap %>% 
  ggplot(aes(x =ANNUAL.INCOME))+
  geom_histogram()

#Desc(cvap$ANNUAL.INCOME)

cvap %>% 
  ggplot(aes(x =ANNUAL.INCOME, color = Loan.Source, fill = Loan.Source))+
  geom_density(alpha =0.1)+theme_bw()

ggplot(data=cvap, mapping=aes(sample=ANNUAL.INCOME)) +
  stat_qq(geom = 'point', distribution = 'qunif') +
  scale_x_continuous(n.breaks = 10) +
  labs(x='Quantile', y='Annual Income', title='Quantile Plot, Annual income')
```


### Loan received distribution

```{r}
cvap %>% 
  ggplot(aes(x =Loan.down.payment))+
  geom_density()+theme_bw()

cvap %>% 
  ggplot(aes(x = LOAN.AMOUNT))+
  geom_density()+theme_bw()

cvap %>% 
  ggplot(aes(x = TOTAL.VEHICLE.COST))+
  geom_density()+theme_bw()

cvap %>% 
  ggplot(aes(x =LOAN.AMOUNT))+
  geom_histogram()



#names(cvap)
```


## Vehicle's profile

### New or Used

```{r}
cvap %>%
  ggplot(aes(x = Purchased.Vehicle.New.or.Used, fill =Purchased.Vehicle.New.or.Used ))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("New vs used")
```

### Tech type vehicle

```{r}
cvap %>%
  ggplot(aes(x = Technology.Type..Fuel.Type, fill =Technology.Type..Fuel.Type))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("TYpe of vehicle")
```

### leased or not

```{r}
cvap %>%
  ggplot(aes(x = Lease..Y.N., fill =Lease..Y.N.))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("Leased (Yes/No)")
```

### How many opt for chargers

```{r}
cvap %>%
  ggplot(aes(x = Charging.Station.Type, fill =Charging.Station.Type))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("Charger received?")

```

### Make of the vehicle

```{r}
make.count <- cvap %>% 
  group_by(MAKE) %>% 
  summarise(count= n())

make.count2 <- make.count %>% 
  mutate(prop = round(count/sum(make.count$count), 2))

make.count1 <- make.count[order(-make.count$count),]

make.count1 %>% 
  ggplot(aes(x = reorder(MAKE, count, sum),  y = count, fill = MAKE))+
  geom_bar(stat = "identity", position = "dodge")+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("Make")+ggtitle( "Count")

make.count2 %>% 
  ggplot(aes(x = reorder(MAKE, prop, sum),  y = prop, fill = MAKE))+
  geom_bar(stat = "identity", position = "dodge")+
  theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
  xlab("Make") + ggtitle("proportion")

# high.prop <- make.count2[make.count2$prop > 0.01,]
# 
# make.count1 <- make.count[order(-make.count$count),]
# cvap %>% 
#   ggplot(aes(MAKE))+
#   geom_bar(fill = "#0073C2FF")+
#   theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
#   xlab("Make")
```

### Vehicle cost

```{r}
cvap %>% 
  ggplot(aes(x =TOTAL.VEHICLE.COST))+
  geom_histogram()
```


```{r}
# cvap[cvap$MAKE %in% high.prop$MAKE,] %>% 
#   ggplot(aes(MAKE))+
#   geom_bar(fill = "#0073C2FF")+
#   theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
#   xlab("Make")+ facet_wrap(.~Loan.Source)
```

```{r}
# cvap %>% 
#   group_by(MAKE, Loan.Source) %>% 
#   summarise(count = n(), avg.laon.amount = mean(LOAN.AMOUNT, na.rm = TRUE))
# 
# cvap[cvap$LOAN.AMOUNT != "No Loan",]%>% 
#   ggplot(aes(x = MAKE, y = LOAN.AMOUNT, fill = Loan.Source))+
#   geom_bar(stat =  "identity", position = "dodge")+
#   theme(axis.text.x = element_text(angle = 60, hjust =1, vjust =1))+
#   xlab("Make")+ facet_wrap(.~Loan.Source)
```


