---
title: "regression_forecast v1_1"
author: "Arun Sharma"
date: "4/24/2021"
output: html_document
---

> Load packages

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(reshape2)

rm(list = ls())
```

# Load datasets

## population
```{r}
#Population
pop <- read.csv("population.csv")

#select only relevant rows
pop <- pop[1:59,1:27]

#remove county from names
pop$Region <- gsub(" County", "", pop$Region)

#remove x from cols
names(pop) <- gsub("X","", names(pop))

long_pop <- melt(pop, id.vars = c("Region"))

names(long_pop)[2] <- "year"
names(long_pop)[3] <- "population"
names(long_pop)[1] <- "county"
long_pop$year <- as.numeric(as.character(long_pop$year))
long_pop$county <- tolower(long_pop$county)
long_pop$county <- gsub("[^a-zA-Z]", "", long_pop$county)
```

## Income
```{r}
#Income
income <- read.csv("forecastedIncome.csv")

income <- income[ ,2:60]

county_names<- names(income)
income$year<- 2000:2035
income <-subset(income, year >=2010)
col_idx <- grep("^year$", names(income))
income <- income[, c(col_idx, (1:ncol(income))[-col_idx])]

income.long <- pivot_longer(income, cols = 2:59, names_to = "county" , values_to= "avg_income")


income.long$county <- tolower(income.long$county)
income.long$county <- gsub("[^a-zA-Z]", "", income.long$county)


county_names <-county_names[-length(county_names)]
county_names <- tolower(county_names)
county_names <- gsub("[^a-zA-Z]", "", county_names)




```

## Charging Stations

```{r}
#Charging Station
chargers <- read.csv("county_chargers.csv")

#charging
chargers1 <- read.csv("county_datapoints.csv")
chargers1$county <- tolower(chargers1$county)
chargers1$county <- gsub("[^a-zA-Z]", "", chargers1$county)

#charger_forecast
charger_forecast <- read.csv("charger_forecast.csv")
names(charger_forecast)[2] <- "chargers"

#chargers forecast
char_for <- read.csv("charger_forecast.csv")
charger_X <-  char_for[char_for$year_list>2020,]
names(charger_X)[2] <- "forecast_charger"
charger_X$county <- tolower(charger_X$county)
charger_X$county <- gsub("[^a-zA-Z]", "", charger_X$county)
charger_forecast<-rename(charger_forecast, c('year'= 'year_list'))
charger_forecast$county <- tolower(charger_forecast$county)
charger_forecast$county <- gsub("[^a-zA-Z]", "", charger_forecast$county)


#combo chargers 
char_all<- read.csv("charger_all.csv")


```


## Fuel price-Gas
```{r}
#Fuel Price
fuel_price <- read.csv("CA_GasPreds_1995_2035.csv")
names(fuel_price) <- c("Date", "GasPrice_gal")
fuel_price$Date <- as.Date(fuel_price$Date, format = "%m/%d/%Y")
fuel_price$year <- year(fuel_price$Date)
f <- fuel_price %>% 
  group_by(year) %>% 
  summarise(avg_price = mean(GasPrice_gal))

gas_price <- f[f$year>2010,]
names(gas_price)[2] <- "gas_price"
```


## Vehicle price ##updated Vehicle price with new CSV
```{r}
#Vehicle price
vehicle_price <- read.csv("Vehicle_prices_final.csv")
names(vehicle_price)[1] <- "year"
```

## Miles per KWH

```{r}
#Miles per kwh
mpk <- read.csv("forecastedMPKwh.csv")
mpk <- mpk[, 2:3]
names(mpk)[1] <- "year"
```
## Y- Number of vehicles sold

```{r}
Y <- read.csv("numEV_CountyandYear.csv")
Y <- Y[, 2:4]

#Y<-Y%>% 
  #group_by(county)%>%
  #mutate(lag.value = lag(num_ev))
Y<-Y%>% 
  group_by(county)%>%
  mutate(lag.value = cumsum(num_ev))

Y$county <- tolower(Y$county)
Y$county <- gsub("[^a-zA-Z]", "", Y$county)
yearly_sales <- Y %>% 
  group_by(year) %>% 
  summarise(total_sales = sum(num_ev, na.rm = TRUE))
```

### Kw prices (All of CA)
```{r}
kw_prices<-read.csv("yearly_kwh_prices.csv")
kw_prices <-subset(kw_prices, year >=2010)
```





## Merging data for Xs and Y

```{r}
final <- merge(Y, gas_price, by = "year")
final <- merge(final, vehicle_price, by = "year")
final <- merge(final, mpk, by = "year")
final <-merge(final, kw_prices, by= "year")
final <- merge(final, long_pop, by = c("county", "year")) 
final<- right_join(final, income.long, by= c("county", "year"))

#lose rows over here
final <- merge(final, chargers1[,c("county","year","csum")], by = c("county", "year"))
#final$avg_income <- (final$low+final$medium+final$high)/3
final<-rename(final, c("chargers" ='csum'))
#model_data<-final %>% group_by(county)%>% 
 # summarise(Avg_County_come= mean(avg_income))
#model_data<-model_data %>% group_by(year)%>% 
#  summarise(Total_year_EV= sum(num_ev))

#head(model_data)
#model_data <- select(final, -c("year", "county", "low", "medium", "high"))

#preping data for model

#scale
```


Xs <- names(model_data)[2:6]
for (i in Xs ){
  print(model_data[,i])
}

scaled_model_data <- model_data
scaled_model_data$gas_price <- c(scale(model_data$gas_price, center = TRUE, scale = TRUE))

scaled_model_data$m_p_k<- c(scale(model_data$m_p_k, center = TRUE, scale = TRUE))

scaled_model_data$population<- c(scale(model_data$population, center = TRUE, scale = TRUE))

scaled_model_data$csum <- c(scale(model_data$csum, center = TRUE, scale = TRUE))

scaled_model_data$avg_income <- c(scale(model_data$avg_income, center = TRUE, scale = TRUE))


## Run the model

```{r}
final<-na.omit(final)

lm0 <- lm(num_ev~.-c(year), data=final)
summary(lm0)

plot(lm0)


lm1 <- lm(num_ev~year, data=final)
summary(lm1)

plot(lm1)
ggplot(final, aes(y=num_ev, x=year+chargers+gas_price+high+low+mid+m_p_k+dollar.per.kw+population+avg_income))+
  stat_summary(fun.data = mean_cl_normal)+
  geom_smooth(method='lm', formula=y~x)


lm2 <- lm(num_ev ~poly(num_ev, x=year+chargers+gas_price+high+low+mid+m_p_k+dollar.per.kw+population+avg_income), data= final)
summary(lm2)

plot(fitted((lm2), residuals(lm2)))



lm3 <- lm(num_ev ~., data= final)
summary(lm3)

```


y_test<- as.data.frame(lm2$fitted.values)
x<- as.data.frame(list(2010:2035))
test<-cbind(x, y_test)

```{r}


char <- na.omit(chargers1)
char<-char[, 2:5]
char$chargers<-char$csum


charger_forecast<- charger_forecast[,-which(names(charger_forecast) %in% c("X"))]
char<- char[,-which(names(char) %in% c("X", "total_chargers", "csum"))]
char<-rbind(char, charger_forecast)
char$county[char$county == "cityandcountyofsanfrancisco"] <- "sanfrancisco"

#char<-char[-grep("nevada", char$county),]


```



##Please note this code is not in a forloop and currently needs to be manually run
```{r}


new_Y<-Y[,-4]
#new_Y <- pivot_wider(new_Y, names_from =year, values_from=num_ev)
#new_Y$'2021'<-0
#new_Y$'2022'<-0
#new_Y$'2023'<-0
#new_Y$'2024'<-0
#new_Y$'2025'<-0
#new_Y$'2026'<-0
#new_Y$'2027'<-0
#new_Y$'2028'<-0
#new_Y$'2029'<-0
#new_Y$'2030'<-0
#new_Y$'2031'<-0
#new_Y$'2032'<-0
#new_Y$'2033'<-0
#new_Y$'2034'<-0
#new_Y$'2035'<-0

#new_Y <- pivot_longer(new_Y, cols = 2:13, names_to = "year", values_to= "num_ev")
new_Y$year<-as.numeric(new_Y$year)
future_x <- merge(new_Y, gas_price, by = "year")
future_x <- merge(future_x, vehicle_price, by = "year")
future_x <- merge(future_x, mpk, by = "year")
future_x <-merge(future_x, kw_prices, by= "year")
future_x <- merge(future_x, long_pop, by = c("county", "year")) 
future_x<- left_join(future_x, income.long, by= c("county", "year"))
future_x <-left_join(future_x, char[,c("county","year","chargers")], by = c("county", "year"))

Y2021<-Y_2021
Y2021$num_ev<- Y_2021$preds
Y2021<- as.data.frame(Y2021[ , -c(13:14)])
future_x<-rbind(future_x, Y2021)

Y2022<-Y_2022
Y2022$num_ev<- Y2022$preds
Y2022<- as.data.frame(Y2022[,-c(13:14)])
future_x<-rbind(future_x, Y2022)

Y2023<-Y_2023
Y2023$num_ev<- Y2023$preds
Y2023<- as.data.frame(Y2023[,-c(13:14)])
future_x<-rbind(future_x, Y2023)

Y2024<-Y_2024
Y2024$num_ev<- Y2024$preds
Y2024<- as.data.frame(Y2024[,-c(13:14)])
future_x<-rbind(future_x, Y2024)

Y2025<-Y_2025
Y2025$num_ev<- Y2025$preds
Y2025<- as.data.frame(Y2025[,-c(13:14)])
future_x<-rbind(future_x, Y2025)

Y2026<-Y_2026
Y2026$num_ev<- Y2026$preds
Y2026<- as.data.frame(Y2026[,-c(13:14)])
future_x<-rbind(future_x, Y2026)

Y2027<-Y_2027
Y2027$num_ev<- Y2027$preds
Y2027<- as.data.frame(Y2027[,-c(13:14)])
future_x<-rbind(future_x, Y2027)

Y2028<-Y_2028
Y2028$num_ev<- Y2028$preds
Y2028<- as.data.frame(Y2028[,-c(13:14)])
future_x<-rbind(future_x, Y2028)

Y2029<-Y_2029
Y2029$num_ev<- Y2029$preds
Y2029<- as.data.frame(Y2029[,-c(13:14)])
future_x<-rbind(future_x, Y2029)

Y2030<-Y_2030
Y2030$num_ev<- Y2030$preds
Y2030<- as.data.frame(Y2030[,-c(13:14)])
future_x<-rbind(future_x, Y2030)

Y2031<-Y_2031
Y2031$num_ev<- Y2031$preds
Y2031<- as.data.frame(Y2031[,-c(13:14)])
future_x<-rbind(future_x, Y2031)

Y2032<-Y_2032
Y2032$num_ev<- Y2032$preds
Y2032<- as.data.frame(Y2032[,-c(13:14)])
future_x<-rbind(future_x, Y2032)

Y2033<-Y_2033
Y2033$num_ev<- Y2033$preds
Y2033<- as.data.frame(Y2033[,-c(13:14)])
future_x<-rbind(future_x, Y2033)

Y2034<-Y_2034
Y2034$num_ev<- Y2034$preds
Y2034<- as.data.frame(Y2034[,-c(13:14)])
future_x<-rbind(future_x, Y2034)

Y2035<-Y_2035
Y2035$num_ev<- Y2035$preds
Y2035<- as.data.frame(Y2035[,-c(13:14)])
future_x<-rbind(future_x, Y2035)

future_x<-future_x%>% 
  group_by(county)%>%
  mutate(lag.value = cumsum(num_ev))

write.csv(future_x, "predictions_CC_bycounty.csv")
```


```{r}
future_x[is.na(future_x)]<-0

#final= subset(future_x, year <=2020)
#final= subset(future_x, year <=2021)
#final= subset(future_x, year <=2022)
#final= subset(future_x, year <=2023)
#final= subset(future_x, year <=2024)
#final= subset(future_x, year <=2025)
#final= subset(future_x, year <=2026)
#final= subset(future_x, year <=2027)
#final= subset(future_x, year <=2028)
#final= subset(future_x, year <=2029)
#final= subset(future_x, year <=2030)
#final= subset(future_x, year <=2031)
#final= subset(future_x, year <=2032)
#final= subset(future_x, year <=2033)
final= subset(future_x, year <=2034)

lm3 <- lm(num_ev ~., data= final)
```


## predict for future

```{r}

prediction <- predict(lm3, future_x, se.fit = TRUE)


preds_cc <- future_x

preds_cc$preds<- prediction$fit

#Y_2021<-subset(preds_cc, year == '2021')  

#Y_2022<-subset(preds_cc, year == '2022')  
#Y_2023<-subset(preds_cc, year == '2023')
#Y_2024<-subset(preds_cc, year == '2024')
#Y_2025<-subset(preds_cc, year == '2025')

#Y_2026<-subset(preds_cc, year == '2026')

#Y_2027<-subset(preds_cc, year == '2027')
#Y_2028<-subset(preds_cc, year == '2028')
#Y_2029<-subset(preds_cc, year == '2029')
#Y_2030<- subset(preds_cc, year == '2030')
#Y_2031<-subset(preds_cc, year == '2031')
#Y_2032<-subset(preds_cc, year == '2032')
#Y_2033<-subset(preds_cc, year == '2033')
#Y_2034<-subset(preds_cc, year == '2034')
Y_2035<-subset(preds_cc, year == '2035')
```


```{r}
ggplot(preds_cc, aes(x=year,y=preds))+ geom_point()

final_preds<- preds_cc%>% group_by(year) %>% summarise(CA_total_ev = sum(preds), Original=sum(num_ev))



```



##Please note this code is not in a forloop and currently needs to be manually run
```{r}

#CA total
new_Y<-Y[,-4]
#new_Y <- pivot_wider(new_Y, names_from =year, values_from=num_ev)
#new_Y$'2021'<-0
#new_Y$'2022'<-0
#new_Y$'2023'<-0
#new_Y$'2024'<-0
#new_Y$'2025'<-0
#new_Y$'2026'<-0
#new_Y$'2027'<-0
#new_Y$'2028'<-0
#new_Y$'2029'<-0
#new_Y$'2030'<-0
#new_Y$'2031'<-0
#new_Y$'2032'<-0
#new_Y$'2033'<-0
#new_Y$'2034'<-0
#new_Y$'2035'<-0

#new_Y <- pivot_longer(new_Y, cols = 2:13, names_to = "year", values_to= "num_ev")
new_Y$year<-as.numeric(new_Y$year)
future_x <- merge(new_Y, gas_price, by = "year")
future_x <- merge(future_x, vehicle_price, by = "year")
future_x <- merge(future_x, mpk, by = "year")
future_x <-merge(future_x, kw_prices, by= "year")
future_x <- merge(future_x, long_pop, by = c("county", "year")) 
future_x<- left_join(future_x, income.long, by= c("county", "year"))
future_x <-left_join(future_x, char[,c("county","year","chargers")], by = c("county", "year"))

future_x[is.na(future_x)]<-0
future_x$avg_car_price<-(future_x$high + future_x$low+ future_x$mid)/3
future_x<- future_x%>% group_by(year)%>% 
  summarise(num_ev= sum(num_ev, na.rm = TRUE), gas_price = mean(gas_price), m_p_k= mean(m_p_k), dollar.per.kw = mean(dollar.per.kw), population=sum(population), avg_income = mean(avg_income),  chargers= sum(chargers), avg_car_price = mean(avg_car_price))


Y2021<-Y_2021
Y2021$num_ev<- Y_2021$preds
Y2021<- as.data.frame(Y2021[ , -c(10:11)])
future_x<-rbind(future_x, Y2021)



Y2022<-Y_2022
Y2022$num_ev<- Y2022$preds
Y2022<- as.data.frame(Y2022[,-c(10:11)])
future_x<-rbind(future_x, Y2022)


Y2023<-Y_2023
Y2023$num_ev<- Y2023$preds
Y2023<- as.data.frame(Y2023[,-c(10:11)])
future_x<-rbind(future_x, Y2023)




Y2024<-Y_2024
Y2024$num_ev<- Y2024$preds
Y2024<- as.data.frame(Y2024[,-c(10:11)])
future_x<-rbind(future_x, Y2024)

Y2025<-Y_2025
Y2025$num_ev<- Y2025$preds
Y2025<- as.data.frame(Y2025[,-c(10:11)])
future_x<-rbind(future_x, Y2025)


Y2026<-Y_2026
Y2026$num_ev<- Y2026$preds
Y2026<- as.data.frame(Y2026[,-c(10:11)])
future_x<-rbind(future_x, Y2026)


Y2027<-Y_2027
Y2027$num_ev<- Y2027$preds
Y2027<- as.data.frame(Y2027[,-c(10:11)])
future_x<-rbind(future_x, Y2027)


Y2028<-Y_2028
Y2028$num_ev<- Y2028$preds
Y2028<- as.data.frame(Y2028[,-c(10:11)])
future_x<-rbind(future_x, Y2028)

Y2029<-Y_2029
Y2029$num_ev<- Y2029$preds
Y2029<- as.data.frame(Y2029[,-c(10:11)])
future_x<-rbind(future_x, Y2029)

Y2030<-Y_2030
Y2030$num_ev<- Y2030$preds
Y2030<- as.data.frame(Y2030[,-c(10:11)])
future_x<-rbind(future_x, Y2030)



Y2031<-Y_2031
Y2031$num_ev<- Y2031$preds
Y2031<- as.data.frame(Y2031[,-c(10:11)])
future_x<-rbind(future_x, Y2031)


Y2032<-Y_2032
Y2032$num_ev<- Y2032$preds
Y2032<- as.data.frame(Y2032[,-c(10:11)])
future_x<-rbind(future_x, Y2032)

Y2033<-Y_2033
Y2033$num_ev<- Y2033$preds
Y2033<- as.data.frame(Y2033[,-c(10:11)])
future_x<-rbind(future_x, Y2033)

Y2034<-Y_2034
Y2034$num_ev<- Y2034$preds
Y2034<- as.data.frame(Y2034[,-c(10:11)])
future_x<-rbind(future_x, Y2034)

Y2035<-Y_2035
Y2035$num_ev<- Y2035$preds
Y2035<- as.data.frame(Y2035[,-c(10:11)])
future_x<-rbind(future_x, Y2035)

write.csv(future_x, "CA_Preds_CC.csv")
```
```{r}
j=as.numeric(future_x[1,2])
future_x$lag.value<-0

for (i in 1:nrow(future_x)){
  future_x[i,10] <- j
  k<-as.numeric(future_x[i,10])
  j=j+k
}




#write.csv(future_x, "predictions_CC.csv")

future_x[is.na(future_x)]<-0

#final= subset(future_x, year <=2020)
#final= subset(future_x, year <=2021)
#final= subset(future_x, year <=2022)
#final= subset(future_x, year <=2023)
#final= subset(future_x, year <=2024)
#final= subset(future_x, year <=2025)
#final= subset(future_x, year <=2026)
#final= subset(future_x, year <=2027)
final= subset(future_x, year <=2028)
#final= subset(future_x, year <=2029)
#final= subset(future_x, year <=2030)
#final= subset(future_x, year <=2031)
#final= subset(future_x, year <=2032)
#final= subset(future_x, year <=2033)
#final= subset(future_x, year <=2034)

lm3 <- lm(num_ev ~., data= final)
```


## predict for future

```{r}

prediction <- predict(lm3, future_x, se.fit = TRUE)


preds_cc <- future_x

preds_cc$preds<- prediction$fit

#Y_2021<-subset(preds_cc, year == '2021')  
#Y_2022<-subset(preds_cc, year == '2022')  
#Y_2023<-subset(preds_cc, year == '2023')
#Y_2024<-subset(preds_cc, year == '2024')
#Y_2025<-subset(preds_cc, year == '2025')
#Y_2026<-subset(preds_cc, year == '2026')
#Y_2027<-subset(preds_cc, year == '2027')
#Y_2028<-subset(preds_cc, year == '2028')
Y_2029<-subset(preds_cc, year == '2029')
#Y_2030<- subset(preds_cc, year == '2030')
#Y_2031<-subset(preds_cc, year == '2031')
#Y_2032<-subset(preds_cc, year == '2032')
#Y_2033<-subset(preds_cc, year == '2033')
#Y_2034<-subset(preds_cc, year == '2034')
#Y_2035<-subset(preds_cc, year == '2035')
```



