---
title: "regression_forecast v1_1"
author: "Arun Sharma"
date: "4/24/2021"
output: html_document
---

> Load packages

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(reshape2)

rm(list = ls())
```

# Load datasets

## population
```{r}
#Population
pop <- read.csv("population.csv")

#select only relevant rows
pop <- pop[1:59,1:27]

#remove county from names
pop$Region <- gsub(" County", "", pop$Region)

#remove x from cols
names(pop) <- gsub("X","", names(pop))

long_pop <- melt(pop, id.vars = c("Region"))

names(long_pop)[2] <- "year"
names(long_pop)[3] <- "population"
names(long_pop)[1] <- "county"
long_pop$year <- as.numeric(as.character(long_pop$year))
long_pop$county <- tolower(long_pop$county)
long_pop$county <- gsub("[^a-zA-Z]", "", long_pop$county)
```

## Income
```{r}
#Income
income <- read.csv("forecastedIncome.csv")
income <- select(income, -X)

county_names<- names(income)
income$year<- 2000:2035
income <-subset(income, year >=2010)
col_idx <- grep("^year$", names(income))
income <- income[, c(col_idx, (1:ncol(income))[-col_idx])]

income.long <- pivot_longer(income, cols = 2:59, names_to = "county" , values_to= "avg_income")


income.long$county <- tolower(income.long$county)
income.long$county <- gsub("[^a-zA-Z]", "", income.long$county)



county_names <-county_names[-length(county_names)]
county_names <- tolower(county_names)
county_names <- gsub("[^a-zA-Z]", "", county_names)




```

## Charging Stations

```{r}
#Charging Station
chargers <- read.csv("county_chargers.csv")

#charging
chargers1 <- read.csv("county_datapoints.csv")
chargers1$county <- tolower(chargers1$county)
chargers1$county <- gsub("[^a-zA-Z]", "", chargers1$county)

#charger_forecast
charger_forecast <- read.csv("charger_forecast.csv")
names(charger_forecast)[2] <- "chargers"

#chargers forecast
char_for <- read.csv("charger_forecast.csv")
charger_X <-  char_for[char_for$year_list>2020,]
names(charger_X)[2] <- "forecast_charger"
charger_X$county <- tolower(charger_X$county)
charger_X$county <- gsub("[^a-zA-Z]", "", charger_X$county)
charger_forecast<-rename(charger_forecast, c('year'= 'year_list'))
charger_forecast$county <- tolower(charger_forecast$county)
charger_forecast$county <- gsub("[^a-zA-Z]", "", charger_forecast$county)
```


## Fuel price-Gas
```{r}
#Fuel Price
fuel_price <- read.csv("CA_GasPreds_1995_2035.csv")
names(fuel_price) <- c("Date", "GasPrice_gal")
fuel_price$Date <- as.Date(fuel_price$Date, format = "%m/%d/%Y")
fuel_price$year <- year(fuel_price$Date)
f <- fuel_price %>% 
  group_by(year) %>% 
  summarise(avg_price = mean(GasPrice_gal))

gas_price <- f[f$year>2010,]
names(gas_price)[2] <- "gas_price"
```


## Vehicle price ##updated Vehicle price with new CSV
```{r}
#Vehicle price
vehicle_price <- read.csv("Vehicle_prices_final.csv")
names(vehicle_price)[1] <- "year"
```

## Miles per KWH

```{r}
#Miles per kwh
mpk <- read.csv("forecastedMPKwh.csv")
mpk <- select(mpk, -X)
names(mpk)[1] <- "year"
```
## Y- Number of vehicles sold

```{r}
Y <- read.csv("numEV_CountyandYear.csv")
Y <- select(Y, -X)

#Y<-Y%>% 
  #group_by(county)%>%
  #mutate(lag.value = lag(num_ev))
Y<-Y%>% 
  group_by(county)%>%
  mutate(lag.value = cumsum(num_ev))

Y$county <- tolower(Y$county)
Y$county <- gsub("[^a-zA-Z]", "", Y$county)
yearly_sales <- Y %>% 
  group_by(year) %>% 
  summarise(total_sales = sum(num_ev, na.rm = TRUE))
```

### Kw prices (All of CA)
```{r}
kw_prices<-read.csv("yearly_kwh_prices.csv")
kw_prices <-subset(kw_prices, year >=2010)
```


## Merging data for Xs and Y

```{r}
final <- merge(Y, gas_price, by = "year")
final <- merge(final, vehicle_price, by = "year")
final <- merge(final, mpk, by = "year")
final <-merge(final, kw_prices, by= "year")
final <- merge(final, long_pop, by = c("county", "year")) 
final<- right_join(final, income.long, by= c("county", "year"))

#lose rows over here
final <- merge(final, chargers1[,c("county","year","csum")], by = c("county", "year"))
#final$avg_income <- (final$low+final$medium+final$high)/3
final<-rename(final, c("chargers" ='csum'))
#model_data<-final %>% group_by(county)%>% 
 # summarise(Avg_County_come= mean(avg_income))
#model_data<-model_data %>% group_by(year)%>% 
#  summarise(Total_year_EV= sum(num_ev))

#head(model_data)
#model_data <- select(final, -c("year", "county", "low", "medium", "high"))

#preping data for model

#scale
```


Xs <- names(model_data)[2:6]
for (i in Xs ){
  print(model_data[,i])
}

scaled_model_data <- model_data
scaled_model_data$gas_price <- c(scale(model_data$gas_price, center = TRUE, scale = TRUE))

scaled_model_data$m_p_k<- c(scale(model_data$m_p_k, center = TRUE, scale = TRUE))

scaled_model_data$population<- c(scale(model_data$population, center = TRUE, scale = TRUE))

scaled_model_data$csum <- c(scale(model_data$csum, center = TRUE, scale = TRUE))

scaled_model_data$avg_income <- c(scale(model_data$avg_income, center = TRUE, scale = TRUE))


## Run the model

```{r}
final<-na.omit(final)

lm0 <- lm(num_ev~.-c(year), data=final)
summary(lm0)

plot(lm0)


lm1 <- lm(num_ev~year, data=final)
summary(lm1)

plot(lm1)
ggplot(final, aes(y=num_ev, x=year+chargers+gas_price+high+low+mid+m_p_k+dollar.per.kw+population+avg_income))+
  stat_summary(fun.data = mean_cl_normal)+
  geom_smooth(method='lm', formula=y~x)


lm2 <- lm(num_ev ~poly(num_ev, x=year+chargers+gas_price+high+low+mid+m_p_k+dollar.per.kw+population+avg_income), data= final)
summary(lm2)

plot(fitted((lm2), residuals(lm2)))



lm3 <- lm(num_ev ~., data= final)
summary(lm3)

```


y_test<- as.data.frame(lm2$fitted.values)
x<- as.data.frame(list(2010:2035))
test<-cbind(x, y_test)


```{r}


new_Y<-Y[,-4]

new_Y <- pivot_wider(new_Y, names_from =year, values_from=num_ev)
new_Y$'2021'<-0
#new_Y$'2022'<-0
#new_Y$'2023'<-0
#new_Y$'2024'<-0
#new_Y$'2025'<-0
#new_Y$'2026'<-0
#new_Y$'2027'<-0
#new_Y$'2028'<-0
#new_Y$'2029'<-0
#new_Y$'2030'<-0
#new_Y$'2031'<-0
#new_Y$'2032'<-0
#new_Y$'2033'<-0
#new_Y$'2034'<-0
#new_Y$'2035'<-0
#new_Y <- pivot_longer(new_Y, cols = 2:27, names_to = "year", values_to= "num_ev")
new_Y <- pivot_longer(new_Y, cols = 2:13, names_to = "year", values_to= "num_ev")
new_Y$year<-as.numeric(new_Y$year)

future_x <- merge(new_Y, gas_price, by = "year")
future_x <- merge(future_x, vehicle_price, by = "year")
future_x <- merge(future_x, mpk, by = "year")
future_x <-merge(future_x, kw_prices, by= "year")
future_x <- merge(future_x, long_pop, by = c("county", "year")) 
future_x<- right_join(future_x, income.long, by= c("county", "year"))
#lose rows over here


future_x <- right_join(future_x, charger_forecast[,c("county","year","chargers")], by = c("county", "year"))
#future_x <- left_join(future_x, Y[, c("lag.value", "county", "year")], by=c("county","year"))
future_x<-future_x%>% 
  group_by(county)%>%
  mutate(lag.value = cumsum(num_ev))
#future_x<-subset(future_x, year > 2020)  

#future_x$num_ev<-0

future_x<- subset(future_x, year <=2021)


future_x<-na.omit(future_x)
```


## predict for future

```{r}

prediction <- predict(lm3, future_x, se.fit = TRUE)

pred.w.plim <- predict(lm3, future_x, interval = "prediction")

pred.w.clim <- predict(lm3,future_x, interval = "confidence")

#matplot(future_x$m_p_k, cbind(pred.w.clim, pred.w.plim[,-1]),
 #       lty = c(1,2,2,3,3), type = "l", ylab = "predicted y")

# length(prediction$fit)

preds_cc <- future_x

preds_cc$preds<- prediction$fit

Y_2021<-subset(preds_cc, year== '2021')  
#Y_2021<-Y_2021$preds & Y_2021$county



ggplot(preds_cc, aes(x=year,y=preds))+ geom_point()

final_preds<- preds_cc%>% group_by(year) %>% summarise(CA_total_ev = sum(preds), Original=sum(num_ev))



```


preds <- as.data.frame(matrix(0,nrow=15, ncol =2))

names(preds) <- c("Year", "projections")

# preds[1,1]



j=1
for (i in 2021:2035){
  preds[j,1] <- i
  j=j+1
}

preds$`projections` <- prediction$fit


##650 down from 800 
```

# use caret
```{r}
library(caret)
preproc <- preProcess(model_data[,c(2:6)], method = c("center", "scale"))
scaled.new <- predict(preproc, newdata = X_pred)

model <- train(num_ev ~ .,
               data = model_data,
               method = "lm",
               preProcess = c('scale', 'center'))
model$modelInfo
predictions <- predict(model, X_pred)
```

```{r}
prediction <- predict(lm2, scaled.new)

summary(prediction)
```

```{r}

write.csv(preds, "Prediction_california.csv")
```

## predictions with model

### get the X's for future


#gas X
gas_price_X <- gas_price[gas_price$year>2020,]

#m_p_k X
mpk_X <- mpk[mpk$year > 2020,]

#vehicle_price X
vehicle_price_X <- vehicle_price[vehicle_price$year > 2020,]

#chargers
c_X <- charger_X %>% 
  group_by(year_list) %>%
  summarise(csum = sum(forecast_charger))
names(c_X)[1] <- "year"

#population
#x_pop <- long_pop[long_pop$year>2020 & long_pop$county == "alameda",] 
x_pop <- long_pop[long_pop$year>2020,]
x_pop<-x_pop%>% group_by(year)%>%
  summarise(population = sum(population, na.rm = TRUE)
            )
#cbind Xs
X_pred <- cbind(gas_price_X[,c("gas_price")],
               m_p_k =mpk_X[,c("m_p_k")], vehicle_price_X[,c("low","medium", "high")],
                c_X[,c("csum")], population = x_pop[,c("population")])

#combining prices for all category

#X_pred$avg_vehicle_price <- (X_pred$low + X_pred$medium + X_pred$high)/3

X_pred <- select(X_pred, -c( "low", "medium", "high"))